---
title: "Prepping the data"
output: html_document
---

```{r}
library(RSocrata)

df_first <- read.socrata(
  "https://data.cityofnewyork.us/resource/ebb7-mvp5.json",
  # app_token = "YOURAPPTOKENHERE",
  # email     = "user@example.com",
  # password  = "fakepassword"
)
```

```{r}
DSNY <- df_first
library(lubridate)
DSNY$month <- lubridate::parse_date_time(DSNY$month, "%y / %m")
```

##Questions to answer:

- How many variables will be worked with?
- How many years worth of waste data is sufficient enough for training, testing and validating sets of data?
- Which boroughs will be studied?
- Which districts will be studied?
- If working with timeseries models, ensure your variable of interest is turned into a ts object. Research the needed frequency and start dates with this data, as we are looking into monthly waste management values.

```{r}
library(tidyverse)
DSNY_second <- DSNY %>% filter(month >= as.Date("2010-01-01") & month <= as.Date("2021-12-01"))
```

Run the code chunk below when you have decided on all the necessary variables that can be created and used on the DSNY data

```{r}
# DSNY_validation <- DSNY %>% filter(month >= as.Date("2021-01-01") & month <= as.Date("2021-12-01"))
```


```{r}
library(tsfknn)
?knn_forecasting
?as.ts()
```

