---
title: "Eighth meeting notes"
author: "Daniel L."
date: "4/08/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide', error=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(RSocrata)
library(tsfknn)
library(fpp3)
library(tsibble)
library(feasts)
library(timetk)
library(modeltime)
library(zoo)
```


## To prepare for this analysis:
- Step 1: Order of differencing
  + TS plot of residuals from ARIMA(0,0,0) w/ constant in not exactly stationary -> d = 1
  
- Step 2: AR() or MA()
1. Obtain residual from current model: ARIMA(0,1,0) w/ constant
2. Plot PACF of the residual

- Step 3: Seasonality()
1. Look at the PACF to determine AR() or MA() terms

```{r echo = FALSE}
#The use of "\\" is also welcomed
load("C:/Users/Daniel/Desktop/R/STAT_790/STAT_790_Case_Seminar/BX_ts.RData")
load("C:/Users/Daniel/Desktop/R/STAT_790/STAT_790_Case_Seminar/sixth_meeting_documents/meeting_6.RDATA")
load("C:\\Users\\Daniel\\Desktop\\R\\STAT_790\\STAT_790_Case_Seminar\\data_prep.Rdata")
#C:\Users\Daniel\Desktop\R\STAT_790\STAT_790_Case_Seminar\sixth_meeting_documents\meeting_6.RData
```


## Creating models with zoo() and the arima package from stats()

Here we create the ts and create the autoplot of the ts using the zoo package

```{r autoplot}
DSNY_BX_zoo_ts <- ts(DSNY_third_bronx[,-1], 
                     start = as.yearmon(DSNY_third_bronx$month)[1], 
                     frequency = 12)

autoplot(as.zoo(DSNY_BX_zoo_ts))
```

### $\text{ARIMA(0,0,0)}$ with constant

```{r arima000}
zoo_arima000_fit <- arima(DSNY_BX_zoo_ts, order = 1 + c(0,0,0))
#names(zoo_arima000_fit)
res_arima000 <- zoo_arima000_fit$residuals
acf(res_arima000, lag.max = 36)
pacf(res_arima000, lag.max = 36)
accuracy(bx_arima000_fit)[4]
```

The lags are in decimal format. With the frequency defined as 12, I believe Lag 1.0 = 12, Lag 2.0 = 24, Lag 3.0 = 36. 

For ARIMA models with differencing, the differenced series follows a zero-mean ARMA model. Documentation by [DataCamp](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/arima).

### $\text{ARIMA(0,1,0)}$ with constant/mean

```{r arima010}
zoo_arima010_fit <- arima(DSNY_BX_zoo_ts, 
                          order = 1 + c(0,1,0)) #the PACF plot return the same values
                          #with or without the include.mean = parameter
res_arima010 <- zoo_arima010_fit$residuals
acf(res_arima010, lag.max = 36)
pacf(res_arima010, lag.max = 36)
accuracy(bx_arima010_fit)[4]
```

A constant term has been added with '1 +  c(0,1,0)'. As discussed before, a negative ACF value is indicative of the need to add an MA() parameter. Here we see that lag 3 is no longer significant, perhaps we can try to add MA(1) and MA(2) arguments to our differenced time series. 

### $\text{ARIMA(0,1,1)}$ with a constant

```{r arima011}
bx_arima011_fit_cons <- bx_ts %>% model(arima011_constant = ARIMA(total_waste ~ 1 + pdq(0,1,1)))
zoo_arima011_fit <- arima(DSNY_BX_zoo_ts, 
                          order = 1 + c(0,1,1))
res_arima011 <- zoo_arima011_fit$residuals
acf(res_arima011, lag.max = 36)
pacf(res_arima011, lag.max = 36)
accuracy(bx_arima011_fit_cons)[4]
```

The lag one value of the ACF plot is 1.0 and positive. With a seasonal pattern of significant lags at each whole lag. There definitely should be a seasonal argument.


```{r arima111, include=FALSE}
### $\text{ARIMA(1,1,1)}$ with constant
# zoo_arima111_fit <- arima(DSNY_BX_zoo_ts, 
#                           order = 1+ c(1,1,1))
# res_arima111 <- zoo_arima111_fit$residuals
# acf(res_arima111, lag.max = 36)
# pacf(res_arima111, lag.max = 36)
#We can see that in the ACF(), the seasonal lags are not being contained within the range. The PACF() is showing promising signs. Lags 4, and 8 are negative and significant and lag 12 is positive and significant. Before committing to a single MA() parameter, I would like to test an MA(2) model on the differenced data. 
```

### $\text{ARIMA(0,1,2)}$ with a constant

```{r arima012}
bx_arima012_cons_fit <- bx_ts %>% model(arima012_cons = ARIMA(total_waste ~ 1 + pdq(0,1,2)))
zoo_arima012_fit <- arima(DSNY_BX_zoo_ts, 
                          order = 1+ c(0,1,2))
res_arima012 <- zoo_arima012_fit$residuals
acf(res_arima012, lag.max = 36)
pacf(res_arima012, lag.max = 36)
accuracy(bx_arima012_cons_fit)[4]
```

Lags (2,4,8) are significant in the PACF(). I am not sure if the lag 2 value of the PACF() is encouraging us to add an AR(2) argument. I would believe that this would remove any progress we have made. For now, attempt an MA(1) model or I think it is best to address the seasonal lags. The RMSE is the lowest so far. 


```{r arima212, include=FALSE}
### $\text{ARIMA(2,1,2)}$ with a constant
# zoo_arima212_fit <- arima(DSNY_BX_zoo_ts, 
#                           order = 1+ c(2,1,2))
# res_arima212 <- zoo_arima212_fit$residuals
# acf(res_arima212, lag.max = 36)
# pacf(res_arima212, lag.max = 36)
#I have attempted an ARIMA(2,1,2) and the PACF() plot is showing promising signs. The non-seasonal lags are contained within range, and only lag 12 is significant. However, I am worried that this model maybe over-fitted. The addition of both AR and MA arguments will make this model hard to interpret.
```


```{r arima212 with seasonal, include=FALSE}
### $\text{ARIMA(2,1,2)}$ with a constant and seasonal period = 12
#$\text{ARIMA(2,1,2)(0,0,0)}_{12}$
# zoo_arima212_000_12fit <- arima(DSNY_BX_zoo_ts, 
#                           order = 1+ c(2,1,2),
#                           seasonal = list(order = c(0L,0L,0L),period = 12))
# res_arima212 <- zoo_arima212_000_12fit$residuals
# acf(res_arima212, lag.max = 36)
# pacf(res_arima212, lag.max = 36)
#It would appear that this 'overfitted' model can not contain the seasonal lags within bounds. For now, I will not focus too much on this model.
```


### $\text{ARIMA(0,1,2)(0,0,1)}$ with a constant and seasonal period = 12
$\text{ARIMA(0,1,2)(0,0,1)}_{12}$

```{r arima012 seasonal}
bx_arima012_cons_001_12 <- bx_ts %>% model(bx_arima012_001_12 = ARIMA(total_waste ~ 1 + 
                                                                   pdq(0,1,2) + 
                                                                   PDQ(0,0,1, period = 12)))
zoo_arima012_001_12fit <- arima(DSNY_BX_zoo_ts, 
                                order = 1 + c(0,1,2),
                                seasonal = list(order = c(0L,0L,1), 
                                                period = 12))
res_arima012_001_12 <- zoo_arima012_001_12fit$residuals
acf(res_arima012_001_12, lag.max = 36)
pacf(res_arima012_001_12, lag.max = 36)
accuracy(bx_arima012_cons_001_12)[4]
```

It appears that the seasonal lags in the PACF() plot appear to still be significant. It also appears that we have bounded the PACF values between (-0.15, 0.2). The RMSE is now the lowest of them all. From meeting 6, I believe that this was the best ARIMA() fitting of them all. But I will try to reduce the MA() parameter by one. 

### $\text{ARIMA(0,1,1)(0,0,1)}$ with a constant and seasonal period = 12
$\text{ARIMA(0,1,1)(0,0,1)}_{12}$

```{r arima011 seasonal}
bx_arima011_cons_001_12 <- bx_ts %>% model(bx_arima011_cons_001_12 = ARIMA(total_waste ~ 1 + 
                                                                             pdq(0,1,1) + 
                                                                             PDQ(0,0,1, period = 12))) 
zoo_arima011_001_12fit <- arima(DSNY_BX_zoo_ts, 
                                order = 1 + c(0,1,1),
                                seasonal = list(order = c(0L,0L,1), 
                                                period = 12))
res_arima011_001_12 <- zoo_arima011_001_12fit$residuals
acf(res_arima011_001_12, lag.max = 36)
pacf(res_arima011_001_12, lag.max = 36)
accuracy(bx_arima011_cons_001_12)[4]
```

In the ACF Plot(), the seasonal lags continue to be significant, as in they pass the 95% threshold. A similar story with the PACF(), but the non-seasonal lags appear to be close to the 95% confidence levels. The PACF values are small as well, so it would not be a major deal breaker. 

### $\text{ARIMA(0,1,1)(1,0,0)}$ with a constant and seasonal period = 12
$\text{ARIMA(0,1,1)(1,0,0)}_{12}$

```{r}
bx_arima011_cons_100_12 <- bx_ts %>% model(bx_arima011_cons_100_12 = ARIMA(total_waste ~ 1 + 
                                                                   pdq(0,1,1) + 
                                                                   PDQ(1,0,0, period = 12)))
zoo_arima011_cons_100_12fit <- arima(DSNY_BX_zoo_ts, 
                                order = 1 + c(0,1,1),
                                seasonal = list(order = c(1,0L,0L), 
                                                period = 12))
res_arima011_cons_100_12 <- zoo_arima011_cons_100_12fit$residuals
acf(res_arima011_cons_100_12, lag.max = 36)
pacf(res_arima011_cons_100_12, lag.max = 36)
accuracy(bx_arima011_cons_100_12)[4]
```

Lowest RMSE

## In conclusion
The models with the most promising ACF and PACF plots are $\text{ARIMA(0,1,1)(1,0,0)}_{12}$, and $\text{ARIMA(0,1,1)(0,0,1)}_{12}$. With my suspicion that the first model is overfitting the timeseries, and the other two models are returning adequate plots. 
