---
title: "R For Data Science: Chapter 19, Functions Practice"
author: "DL"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df <- tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

```{r}
x <- df$a
rng <- range(x, na.rm = TRUE)
(x - rng[1])/(rng[2] - rng[1])
```
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1])/(rng[2] - rng[1])
}

rescale01(c(0,3,9))
```

```{r}
d <- df$b 
rescale01(d)
```

```{r}
df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)
```

```{r}
x <- c(1:10, Inf); x
rescale01(x)
```

Because we’ve extracted the code into a function, we only need to make the fix in one place:

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(x)
```

## 19.2.1 Exercises

1. Why is TRUE not a parameter to rescale01()? What would happen if x contained a single missing value, and na.rm was FALSE?

```{r}
rescale01_alt <- function(x, na.rm = FALSE) {
  rng <- range(x, na.rm = na.rm, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01_alt(c(NA, 1:5), na.rm = FALSE)
```

The option finite = TRUE to range() will drop all non-finite elements, and NA is a non-finite element.

However, if both finite = FALSE and na.rm = FALSE, then this function will return a vector of NA values. Recall, arithmetic operations involving NA values return NA.

```{r}
rescale02_alt2 <- function(x, na.rm = FALSE, finite = FALSE) {
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}

rescale02_alt(c(NA, 1:6), na.rm = FALSE, finite = FALSE)
```

2. In the second variant of rescale01(), infinite values are left unchanged. Rewrite rescale01() so that -Inf is mapped to 0, and Inf is mapped to 1.

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  y <- (x - rng[1]) / (rng[2] - rng[1]) #create this as a variable
  y[y == -Inf] <- 0 #set these values as 0
  y[y == Inf] <- 1 #set these values as 1
  y
}

rescale01(c(Inf, -Inf, 0:5, NA))
```

