---
title: "R For Data Science: Chapter 19, Functions Practice"
author: "DL"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
df <- tibble::tibble(
  a = rnorm(10),
  b = rnorm(10),
  c = rnorm(10),
  d = rnorm(10)
)
```

```{r}
x <- df$a
rng <- range(x, na.rm = TRUE)
(x - rng[1])/(rng[2] - rng[1])
```
```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1])/(rng[2] - rng[1])
}

rescale01(c(0,3,9))
```

```{r}
d <- df$b 
rescale01(d)
```

```{r}
df$a <- rescale01(df$a)
df$b <- rescale01(df$b)
df$c <- rescale01(df$c)
df$d <- rescale01(df$d)
```

```{r}
x <- c(1:10, Inf); x
rescale01(x)
```

Because we’ve extracted the code into a function, we only need to make the fix in one place:

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01(x)
```

## 19.2.1 Exercises

#### 1. Why is TRUE not a parameter to rescale01()? What would happen if x contained a single missing value, and na.rm was FALSE?

```{r}
rescale01_alt <- function(x, na.rm = FALSE) {
  rng <- range(x, na.rm = na.rm, finite = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale01_alt(c(NA, 1:5), na.rm = FALSE)
```

The option finite = TRUE to range() will drop all non-finite elements, and NA is a non-finite element.

However, if both finite = FALSE and na.rm = FALSE, then this function will return a vector of NA values. Recall, arithmetic operations involving NA values return NA.

```{r}
rescale02_alt2 <- function(x, na.rm = FALSE, finite = FALSE) {
  rng <- range(x)
  (x - rng[1]) / (rng[2] - rng[1])
}

rescale02_alt2(c(NA, 1:6), na.rm = FALSE, finite = FALSE)
```

#### 2. In the second variant of rescale01(), infinite values are left unchanged. Rewrite rescale01() so that -Inf is mapped to 0, and Inf is mapped to 1.

```{r}
rescale01 <- function(x) {
  rng <- range(x, na.rm = TRUE, finite = TRUE)
  y <- (x - rng[1]) / (rng[2] - rng[1]) #create this as a variable
  y[y == -Inf] <- 0 #set these values as 0
  y[y == Inf] <- 1 #set these values as 1
  y
}

rescale01(c(Inf, -Inf, 0:5, NA))
```

#### 3. Practice turning the following code snippets into functions. Think about what each function does. What would you call it? How many arguments does it need? Can you rewrite it to be more expressive or less duplicative?

mean(is.na(x)) - this code calculates the proportion of NA values in a vector

```{r}
#We want to return the mean of a vector? The is.na(x) will return T or F for individual elements of this vector?

is.na(c(0, 1, 2, NA, 4, NA)) #for example

prop_na <- function(x){#proportion_na is the name & one input
  mean(is.na(x))
}

print(prop_na(c(11,22,33, NA,NA,NA))) # there are 3 NA's and 3 numeric values, hence the proportion is 0.5: 3/6 values in the vector are NA
```

x / sum(x, na.rm = TRUE) - this will standardize every element in the vector

* set na.rm = FALSE by default in order to make the function behave the same as the built-in functions like sum() and mean() in its handling of missing values.

```{r}
standardize_vector <- function(x, na.rm = FALSE){ #na.rm will be the optional parameter
  return(x/sum(x, na.rm = na.rm)) #return was not necessary
}
```

```{r}
sum_to_one <- function(x, na.rm = FALSE) {
  x / sum(x, na.rm = na.rm)
}
```


```{r}
standardize_vector(1:5)
#sum_to_one(1:5)
standardize_vector(c(1:5,NA)) # drop missing values when standardizing
standardize_vector(c(1:5,NA), na.rm = TRUE)
```

sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE) - This code calculates th coefficient of variation (CV), and is defined as the ratio of the standard deviation $\sigma$ to $\mu$}

```{r}
coef_variation <- function(x, na.rm = FALSE){
  return(sd(x, na.rm = na.rm)/mean(x, na.rm = na.rm))
}
```

```{r}
coef_variation(c(1:10))
coef_variation(c(1:5,NA))
coef_variation(c(1:5,NA), na.rm = TRUE)
```

#### 4.write your own functions to compute the variance and skewness of a numeric vector. Variance is defined as

$$
Var(x)=\frac{1}{n-1} \sum_{i=1}^n (x_i - \bar{x}) ^2 \text{,}
$$
where $\bar{x} = (\sum_i^n x_i) / n$ is the sample mean. Skewness is defined as

$$
\mathrm{Skew}(x) = \frac{\frac{1}{n-2}\left(\sum_{i=1}^n(x_i - \bar x)^3\right)}{\mathrm{Var}(x)^{3/2}} \text{.}
$$

```{r}
variance <- function(x, na.rm = TRUE){
  n = length(x)
  xbar <- mean(x, na.rm = TRUE)
  squared_errors <- (x - xbar)^2
  return(sum(squared_errors/(n-1)))
}
```

```{r}
variance(c(1:10))
```

```{r}
skew <- function(x, na.rm = TRUE){
  n = length(x)
  xbar <- mean(x, na.rm = TRUE)
  denom <- var(x)
  #return(((1/(n-2))*(sum((x-xbar)^3)))/denom^(3/2))
  return((sum((x - xbar) ^ 3) / (n - 2)) / denom ^ (3 / 2))
}
```

```{r}
skew(c(1,3,100))
```

#### 5. Write both_na(), a function that takes two vectors of the same length and returns the number of positions that have an NA in both vectors.

Do we return False if they are not the same length?

```{r}

```

#### 6. Create a function what will return TRUE if a given integer is inside a vector.

```{r}
f_exists <- function(v,x, na.rm = TRUE){ #v is the vector and x is the integer to check
  exist <- FALSE
  i <- 1 #begin the cursor at 1
  while (i <= length(v) & !exist){ #while loop that goes through each element and !exist is TRUE
    if (v[i] == x){ #element wise checking
      exist <- TRUE #exist is switched to TRUE
    }
  i <- i + 1
  }
  return (exist)
}

f_exists(c(33,12,22), 11)
```

#### 7. Create a function that given a data frame will print by screen the name of the column and the class of data it contains (e.g. Variable1 is Numeric).

```{r}
#we need a df parameter, 
f_class <- function(df){
  # a for loop through each column of the df
  for (i in 1:ncol(df)){
    # concatenate the name of the column, with its class type
    cat(names(df)[i], "is", class(df[,i]), "\n")
  }
}
f_class(cars)
```

