---
title: "DOE_data"
author: "DL"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## NYC DOE School Enrollment

Link for Open Data: https://data.cityofnewyork.us/browse?q=Enrollment%20Statistics%20by%20District&sortBy=relevance


```{r}
#https://dev.socrata.com/foundry/data.cityofnewyork.us/e649-r223
doe_first <- read.socrata(
  "https://data.cityofnewyork.us/resource/e649-r223.json"
  # app_token = "YOURAPPTOKENHERE",
  # email     = "user@example.com",
  # password  = "fakepassword"
)
```

```{r}
glimpse(doe_first)
```

We dont necessarily need to convert the year into a posixct. This represents a school year. I.e 2020-2021

My next goal is to explore how column sum or row sum will work on this data set. But first we need to convert the grades from characters to integers

```{r}
#doe_first <- doe_first_copy
doe_first_copy <- doe_first
doe_first[,4:19] <- lapply(doe_first[,4:19], as.integer)

#Line 90 will convert all the NA's in the integer columns into zeros
#doe_first %>% mutate_if(is.integer, function(x){ifelse(is.na(x), 0,x)})
# doe_first %>% 
#   summarise(across(doe_first[,4:19]), ~sum(., is.na(.), 0))

```
```{r}
doe_first %>% 
  mutate(across(doe_first[, 4:19], ~ sum(., is.na(.), 0)))
```

```{r}
doe_first %>% summarise(across(where(is.numeric), ~ sum(.x)))
```

```{r}
doe_first <- doe_first %>% 
  mutate(Total = select(., pk:se1) %>% rowSums(na.rm = TRUE))
```


```{r}
# doe_first %>% 
#   mutate(total_sum = summarise(across(where(is.numeric), ~ sum(.x))))
```

```{r}
glimpse(doe_first)
```

## Another attempt at the NYC DOE Data

```{r}
#https://dev.socrata.com/foundry/data.cityofnewyork.us/xzy8-qqgf
doe_second <- read.socrata("https://data.cityofnewyork.us/resource/xzy8-qqgf.csv")
```

```{r}
glimpse(doe_second)
```

```{r}
which(sapply(doe_second, function(x) is.character(x)))
```

I don't want to make the 'year' as an datetime just yet. But all of these 

```{r}
doe2_col_nums <- c(4:18,20)
```

```{r}
#df_first[cols_num_ints] <- sapply(df_first[cols_num_ints], as.integer)
doe_second[doe2_col_nums] <- sapply(doe_second[doe2_col_nums], as.numeric)
#glimpse(doe_second)
```