---
title: "Queens Time Series"
author: "Daniel L."
date: "5/4/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide', error=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(RSocrata)
library(tsfknn)
library(fpp3)
library(tsibble)
library(tseries)
library(feasts)
library(timetk)
library(modeltime)
library(zoo)
load("C:\\Users\\Daniel\\Desktop\\R\\STAT_790\\STAT_790_Case_Seminar\\Queens_time_series\\QNS_ts_data.Rdata")
```

```{r}
qns_ts %>% 
  ggplot(mapping = aes(x = (month),
                       y = total_waste)) + 
  geom_line() + 
  geom_point(color = "steel blue") + 
  labs(x = "Year",
       y = "Total Waste",
       title = "Queens Total Waste Collected",
       subtitle = "01/2005 - 12/2020")
```

#### seasonal plots

```{r}
qns_ts2 %>% 
  gg_season(total_waste, period = "year", labels = "right") + labs(title = "Seasonal Plot: Total Waste Collected in Queens", x = "Month", y = "Total Waste")

qns_ts2 %>% 
  gg_subseries(total_waste) +
  labs(y = "Total Waste",
       title = "Seasonal Plot: Total Waste Collected in Queens")
```

```{r}
DSNY_third_queens %>% 
  group_by("Year" = year(month)) %>% 
  summarise(., 
            "Average Total Waste" = mean(total_waste))
```

#### KPSS Test for 'total_waste'

$H_0: \text{The time series is trend stationary}$ vs $H_a: \text{The time series is not trend stationary}$

If the p-value of the test is less than some significance level (e.g. $\alpha = .05$) then we reject the null hypothesis and conclude that the time series is not trend stationary.

```{r}
#total waste values
qns_ts %>% features(total_waste, unitroot_kpss)

#differenced values
qns_ts %>% features(diff1, unitroot_kpss)
```

According the KPSS test, both the QNS total_waste and differenced values fail to reject the $H_0$. 

# Begin by looking at ACF and PACF of the total_waste and differenced values

```{r}
qns_ts3 %>% 
  ACF(total_waste, lag_max = 36) %>% 
  autoplot()

#acf of the differenced values
qns_ts3 %>% 
  ACF(diff1, lag_max = 36) %>% 
  autoplot()

qns_ts3 %>% 
  ggplot(mapping = aes(x = month, y = diff1)) + geom_line() +
  labs(x = "Year", 
       y = "Total Waste", 
       title = "Differenced Values: Queens Total Waste Collected")
```

# Creating models with zoo() and the arima package from stats()

```{r}
DSNY_QNS_zoo_ts <- ts(DSNY_third_queens[,2],
   start = as.yearmon(DSNY_third_queens$month)[1],
   frequency = 12)
```

## $\text{ARIMA(0,0,0)}$ with constant
${ARIMA(0,0,0)}$

```{r}
qns_arima000_fit_cons <- qns_ts3 %>% 
  model(arima000_constant = ARIMA(total_waste ~ 1 + pdq(0,0,0)))

zoo_arima000_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(0,0,0))
res_arima000 <- zoo_arima000_fit$residuals
acf(res_arima000, lag.max = 36)
pacf(res_arima000, lag.max = 36)
accuracy(qns_arima000_fit_cons)[4]
```

RMSE = 7382.682. In the ACF and PACF, the first significant lag is lag = 4 and has a negative autocorrelation value. 

Before I add MA(4), I would like to see how the plots of a model with a seasonal AR(1)

## $\text{ARIMA(0,0,0)(1,0,0)[12]}$ with constant
${ARIMA(0,0,0)}(1,0,0)[12]$

```{r}
qns_arima000_fit_seasonal_cons <- qns_ts3 %>% 
  model(arima000_constant_seasonal = ARIMA(total_waste ~ 1 + 
                                     pdq(0,0,0) + 
                                     PDQ(1,0,0, period = 12)))
        
zoo_arima000_seasonal_fit <- arima(DSNY_QNS_zoo_ts, 
                          order = 1 + c(0,0,0),
                          seasonal = list(order = c(1,0L,0L), period = 12))
#names(zoo_arima000_fit)
res_arima000_seasonal <- zoo_arima000_seasonal_fit$residuals
acf(res_arima000_seasonal, lag.max = 36)
pacf(res_arima000_seasonal, lag.max = 36)
accuracy(qns_arima000_fit_seasonal_cons)[4]
```

RMSE = 4413.677. The seasonal lags appear to be contained. The PACF plot shows us that lag = 3 has positive correlation. Lags = (11,23,35) also have positive correlation. 

## $\text{ARIMA(3,0,0)(1,0,0)[12]}$ with constant
${ARIMA(3,0,0)}(1,0,0)[12]$

```{r}
qns_arima300_fit_seasonal_cons <- qns_ts3 %>% 
  model(arima300_constant_seasonal = ARIMA(total_waste ~ 1 + 
                                     pdq(3,0,0) + 
                                     PDQ(1,0,0, period = 12)))
        
zoo_arima300_seasonal_fit <- arima(DSNY_QNS_zoo_ts, 
                          order = 1 + c(3,0,0),
                          seasonal = list(order = c(1,0L,0L), period = 12))
#names(zoo_arima000_fit)
res_arima300_seasonal <- zoo_arima300_seasonal_fit$residuals
acf(res_arima300_seasonal, lag.max = 36)
pacf(res_arima300_seasonal, lag.max = 36)
accuracy(qns_arima300_fit_seasonal_cons)[4]
qns_arima300_fit_seasonal_cons %>% report()
```

RMSE = 4115.02. We do see a decrease in the RMSE. The first 11 lags in the PACF have low autocorrelations. Lag = 12 has significant negative autocorrelation. Lags = (11, 23, 35) continue to be significant.

Lets work with an MA(4) model, with no seasonality. 

## $\text{ARIMA(0,0,4)}$ with constant
${ARIMA(0,0,4)}$

```{r}
qns_arima004_fit_cons <- qns_ts3 %>% 
  model(arima004_constant = ARIMA(total_waste ~ 1 + pdq(0,0,4)))

zoo_arima004_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(0,0,4))
res_arima004 <- zoo_arima004_fit$residuals
acf(res_arima004, lag.max = 36)
pacf(res_arima004, lag.max = 36)
accuracy(qns_arima004_fit_cons)[4]
```

Our RMSE = 5657.673 is a decrease from the first model. Most of the autocorrelations in the ACF plot appear to be contained The first significant lag in the PACF plot is lag = 9, and has negative autocorrelation. Lag = 12 is very positively auto correlated. 

Lets work with an MA(4) model, with seasonality. 

## $\text{ARIMA(0,0,4)(1,0,0)[12]}$ with constant
${ARIMA(0,0,4)(1,0,0)[12]}$

```{r}
qns_arima004_100_fit_cons <- qns_ts3 %>% 
  model(arima004_100_constant = ARIMA(total_waste ~ 1 + 
                                    pdq(0,0,4) +
                                    PDQ(1,0,0, period = 12)))

zoo_arima004_100_fit <- arima(DSNY_QNS_zoo_ts, 
                              order = 1 + c(0,0,4),
                              seasonal = list(order = c(1, 0L, 0L),
                                              period = 12))

res_arima004_100 <- zoo_arima004_100_fit$residuals
acf(res_arima004_100, lag.max = 36)
pacf(res_arima004_100, lag.max = 36)
accuracy(qns_arima004_100_fit_cons)[4]
```

RMSE = 4162.64. The first 10 lags are not significant in the PACF plot. The seasonal lags are negatively autocorrelated. Other lags are also significant. This model does not appear to produce a stable model, according to their plots. 

## $\text{ARIMA(0,1,0)}$ with constant
${ARIMA(0,1,0)}$

```{r}
qns_arima010_fit_cons <- qns_ts3 %>% 
  model(arima010_constant = ARIMA(total_waste ~ 1 + pdq(0,1,0)))

zoo_arima010_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(0,1,0))
res_arima010 <- zoo_arima010_fit$residuals
acf(res_arima010, lag.max = 36)
pacf(res_arima010, lag.max = 36)
accuracy(qns_arima010_fit_cons)[4]
```

RMSE = 6502.128. In both the ACF and PACF plot, we do see a tiny lag-1 negative autocorrelation. According to rule 7, "The lag at which the ACF cuts off is the indicated number of MA terms." However, we also see that the first significant lag is lag-4, which is also negative. Lag-12 is positively correlated and significant. 

Let's explore a MA(1) model

## $\text{ARIMA(0,1,1)}$ with constant
${ARIMA(0,1,1)}$

```{r}
qns_arima011_fit_cons <- qns_ts3 %>% 
  model(arima011_constant = ARIMA(total_waste ~ 1 + pdq(0,1,1)))

zoo_arima011_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(0,1,1))
res_arima011 <- zoo_arima011_fit$residuals
acf(res_arima011, lag.max = 36)
pacf(res_arima011, lag.max = 36)
accuracy(qns_arima011_fit_cons)[4]
qns_arima011_fit_cons %>% report()
```

RMSE = 6421.289, which is a decrease from the base differenced model. The lags = (1,2,3) are still not significant and we see that lag=1 in the PACF has become positive. What we need to keep in mind is that the seasonal lags, lag = 12 is approximately 0.5. Its possible to add a seasonal difference, or add a seasonal AR(1). 

## $\text{ARIMA(0,1,1)(1,0,0)}$ with constant and seasonality

${ARIMA(0,1,1)(1,0,0)[12]}$

```{r}
qns_arima011_100_fit_cons <- qns_ts3 %>% 
  model(arima011_100_constant = ARIMA(total_waste ~ 1 + 
                                        pdq(0,1,1) + 
                                        PDQ(1,0,0,
                                            period = 12)))

zoo_arima011_100_fit <- arima(DSNY_QNS_zoo_ts, 
                              order = 1 + c(0,1,1),
                              seasonal = list(order = c(1,0L,0L),
                                              period = 12))
res_arima011_100 <- zoo_arima011_100_fit$residuals
acf(res_arima011_100, lag.max = 36)
pacf(res_arima011_100, lag.max = 36)
accuracy(qns_arima011_100_fit_cons)[4]
qns_arima011_100_fit_cons %>% report()
```

A decrease in the RMSE = 4174.407. Now lag = 3 is significant and positively correlated. Lags = 11, 23, 35 are positive and significant. We also see that lag = 12, the seasonal lag is also significant. Although we see a decrease in the RMSE, the plots of PACF show mixing autocorrelations at high order lags.

## $\text{ARIMA(3,1,1)(1,0,0)}$ with constant and seasonality

${ARIMA(3,1,1)(1,0,0)[12]}$

```{r}
qns_arima311_100_fit_cons <- qns_ts3 %>% 
  model(arima311_100_constant = ARIMA(total_waste ~ 1 + 
                                        pdq(3,1,1) + 
                                        PDQ(1,0,0,
                                            period = 12)))

zoo_arima311_100_fit <- arima(DSNY_QNS_zoo_ts, 
                              order = 1 + c(3,1,1),
                              seasonal = list(order = c(1,0L,0L),
                                              period = 12))
res_arima311_100 <- zoo_arima311_100_fit$residuals
acf(res_arima311_100, lag.max = 36)
pacf(res_arima311_100, lag.max = 36)
accuracy(qns_arima311_100_fit_cons)[4]
qns_arima311_100_fit_cons %>% report()
```

## $\text{ARIMA(0,1,4)}$ with constant

${ARIMA(0,1,4)}$

```{r}
qns_arima014_fit_cons <- qns_ts3 %>% 
  model(arima014_constant = ARIMA(total_waste ~ 1 + pdq(0,1,4)))

zoo_arima014_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(0,1,4))
res_arima014 <- zoo_arima014_fit$residuals
acf(res_arima014, lag.max = 36)
pacf(res_arima014, lag.max = 36)
accuracy(qns_arima014_fit_cons)[4]
```

RMSE = 5649.113

## $\text{ARIMA(0,1,4)(1,0,0)[12]}$ with constant and seasonality

${ARIMA(0,1,4)(1,0,0)[12]}$

```{r}
qns_arima014_100_fit_cons <- qns_ts3 %>% 
  model(arima014_100_constant = ARIMA(total_waste ~ 1 + 
                                        pdq(0,1,4) + 
                                        PDQ(1,0,0,
                                            period = 12)))

zoo_arima014_100_fit <- arima(DSNY_QNS_zoo_ts, 
                              order = 1 + c(0,1,4),
                              seasonal = list(order = c(1,0L,0L),
                                              period = 12))
res_arima014_100 <- zoo_arima014_100_fit$residuals
acf(res_arima014_100, lag.max = 36)
pacf(res_arima014_100, lag.max = 36)
accuracy(qns_arima014_100_fit_cons)[4]
```

RMSE = 4123.31. Bounded b/w (-0.2 and 0.2)

## AutoArima

```{r}
qns_auto_arima_fit_cons <- qns_ts3 %>% 
  model(stepwise = ARIMA(total_waste),
        search = ARIMA(total_waste, 
                       stepwise = FALSE, 
                       approximation = FALSE))
accuracy(qns_auto_arima_fit_cons)[1:4]
```

```{r}
qns_auto_arima_fit_cons %>% select(.model = stepwise) %>% report()
print("--------------")
qns_auto_arima_fit_cons %>% select(.model = search) %>% report()
```

```{r}
qns_arima302_fit_cons <- qns_ts3 %>% 
  model(arima302_constant = ARIMA(total_waste ~ 1 + pdq(3,0,2)))

zoo_arima302_fit <- arima(DSNY_QNS_zoo_ts, order = 1 + c(3,0,2))
res_arima302 <- zoo_arima302_fit$residuals
acf(res_arima302, lag.max = 36)
pacf(res_arima302, lag.max = 36)
accuracy(qns_arima302_fit_cons)[4]
```

## Summary of models

${ARIMA(0,0,0)}(1,0,0)[12]$ has $RMSE \approx 4413.677$ \
${ARIMA(3,0,0)}(1,0,0)[12]$ has $RMSE \approx 4115.02$ and $AICc \approx 3766.23$ \
${ARIMA(0,0,4)(1,0,0)[12]}$ has $RMSE \approx 4162.64$ \
${ARIMA(3,1,1)(1,0,0)[12]}$ has $RMSE \approx 4111.87$ and $AICc \approx 3750.19$ \
$ARIMA(3,0,2)$ has $RMSE \approx 5477.051$ and $AICc \approx 3866.63$
