---
title: "Brooklyn Total Waste"
author: "Daniel L."
date: "5/2/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, results='hide', error=FALSE, include=FALSE}
library(tidyverse)
library(lubridate)
library(RSocrata)
library(tsfknn)
library(fpp3)
library(tsibble)
library(tseries)
library(feasts)
library(timetk)
library(modeltime)
library(zoo)
load("C:\\Users\\Daniel\\Desktop\\R\\STAT_790\\STAT_790_Case_Seminar\\BK_ts_data.Rdata")
```

```{r}
(bk_ts) %>% 
  ggplot(mapping = aes(x = (month),
                       y = total_waste)) + 
  geom_line() + 
  labs(x = "Year",
       y = "Total Waste",
       title = "Brooklyn Total Waste Collected",
       subtitle = "01/2005 - 12/2020")
```

The majority of the tonnage values are bounded b/w (65000,90000).

#### KPSS Test for 'total_waste'
\
$H_0: \text{The time series is trend stationary}$ vs $H_a: \text{The time series is not trend stationary}$

If the p-value of the test is less than some significance level (e.g. $\alpha = .05$) then we reject the null hypothesis and conclude that the time series is not trend stationary.

```{r kpss test}
#total waste values
bk_ts %>% features(total_waste, unitroot_kpss)
#differenced values
bk_ts %>% features(diff1, unitroot_kpss)
```

Oddly enough, the kpss test on the total waste values has us failing to reject $H_0$. The total waste values are trend stationary. The differenced values are also trend stationary, according the results of the kpss test.

```{r echo=FALSE, results='hide', error=FALSE, include=FALSE}
kpss.test(bk_ts$total_waste, null = "Trend")
```

# Begin by looking at ACF and PACF of the total_waste and differenced values

```{r}
bk_ts3 %>% 
  ACF(total_waste, lag_max = 36) %>% 
  autoplot()

#acf of the differenced values
bk_ts3 %>% 
  ACF(diff1, lag_max = 36) %>% 
  autoplot()

bk_ts3 %>% 
  ggplot(mapping = aes(x = month, y = diff1)) + geom_line() +
  labs(x = "Year", y = "Total Waste", title = "Differenced Values: Brooklyn Total Waste Collected")
```

# Creating models with zoo() and the arima package from stats()

```{r}
DSNY_BK_zoo_ts <- ts(DSNY_third_brooklyn[,2],
   start = as.yearmon(DSNY_third_brooklyn$month)[1],
   frequency = 12)

autoplot(as.zoo(DSNY_BK_zoo_ts))
```

## $\text{ARIMA(0,0,0)}$ with constant
$\text{ARIMA(0,0,0)}$

```{r}
bk_arima000_fit_cons <- bk_ts3 %>% 
  model(arima000_constant = ARIMA(total_waste ~ 1 + pdq(0,0,0)))

zoo_arima000_fit <- arima(DSNY_BK_zoo_ts, order = 1 + c(0,0,0))
res_arima000 <- zoo_arima000_fit$residuals
acf(res_arima000, lag.max = 36)
pacf(res_arima000, lag.max = 36)
accuracy(bk_arima000_fit_cons)[4]
```

We begin with an RMSE = 6036.26. In the ACF plot, the lags that stand out, are positive and significant are lags = (12,24,36). In the PACF plot, the first significant lag is lag = 4, which is negative. Oddly enough, lag = 12 is only seasonal lag that is positive and significant.

Before working with the differenced values, I will like to try different p,d,q arguments. Off the back, I would like to try adding a q = 4. But I will first try working with the seasonal parameters, at an attempt to see if the RMSE decreases

## $\text{ARIMA(0,0,0)(1,0,0)}$ with constant and seasonal parameter
$\text{ARIMA(0,0,0)(1,0,0)}$

```{r}
bk_arima000_fit_seasonal_cons <- bk_ts3 %>% 
  model(arima000_constant_seasonal = ARIMA(total_waste ~ 1 + 
                                     pdq(0,0,0) + 
                                     PDQ(1,0,0, period = 12)))
        
zoo_arima000_seasonal_fit <- arima(DSNY_BK_zoo_ts, 
                          order = 1 + c(0,0,0),
                          seasonal = list(order = c(1,0L,0L), period = 12))
#names(zoo_arima000_fit)
res_arima000_seasonal <- zoo_arima000_seasonal_fit$residuals
acf(res_arima000_seasonal, lag.max = 36)
pacf(res_arima000_seasonal, lag.max = 36)
accuracy(bk_arima000_fit_seasonal_cons)[4]
```

RMSE = 4589.42. The majority of the lags in the ACF plot are contained within bounds. Along with the lags of the PACF plot. Only lag = 11 is significant and positive. The values are bounded b/w (-0.15,0.15).

## $\text{ARIMA(0,0,4)}$ with constant
$\text{ARIMA(0,0,4)}$

```{r}
bk_arima004_fit_cons <- bk_ts3 %>% 
  model(arima004_constant = ARIMA(total_waste ~ 1 + 
                                    pdq(0,0,4)))
        
zoo_arima004_fit <- arima(DSNY_BK_zoo_ts, 
                          order = 1 + c(0,0,4))
#names(zoo_arima000_fit)
res_arima004 <- zoo_arima004_fit$residuals
acf(res_arima004, lag.max = 36)
pacf(res_arima004, lag.max = 36)
accuracy(bk_arima004_fit_cons)[4]
```

RMSE = 5325.188. We do see a decrease in the RMSE when compared to the first $\text{ARIMA(0,0,0)}$. In both the ACF and PACF plots, the seasonal lags are significant.

Lets work with this model and add a seasonal argument.

## $\text{ARIMA(0,0,4)(1,0,0)}$ with constant and seasonal
$ARIMA(0,0,4)(1,0,0)_{12}$
```{r}
bk_arima004_100_seasonal_fit_cons <- bk_ts3 %>% 
  model(arima004_constant = ARIMA(total_waste ~ 1 + 
                                    pdq(0,0,4) + 
                                    PDQ(1,0,0, 
                                        period = 12)))
        
zoo_arima004_100_seasonalfit <- arima(DSNY_BK_zoo_ts, 
                          order = 1 + c(0,0,4),
                          seasonal = list(order = c(1, 0L, 0L), period = 12))
#names(zoo_arima000_fit)
res_arima004_100 <- zoo_arima004_100_seasonalfit$residuals
acf(res_arima004_100, lag.max = 36)
pacf(res_arima004_100, lag.max = 36)
accuracy(bk_arima004_100_seasonal_fit_cons)[4]
```

RMSE = 4444.908. Most of the lagged values are not-significant. They look good and appear to be bounded.

Since we can also work with the differenced values, we will create some models with them

## ${ARIMA(0,1,0)}$ with constant

${ARIMA(0,1,0)}$

```{r}
bk_arima010_fit_cons <- bk_ts3 %>% 
  model(arima010_constant = ARIMA(total_waste ~ 1 + pdq(0,1,0)))

zoo_arima010_fit <- arima(DSNY_BK_zoo_ts, order = 1 + c(0,1,0))
res_arima010 <- zoo_arima010_fit$residuals
acf(res_arima010, lag.max = 36)
pacf(res_arima010, lag.max = 36)
accuracy(bk_arima010_fit_cons)[4]
```

RMSE = 6628.688. Try q = 2 or q = 4

## ${ARIMA(0,1,2)}$ with constant
${ARIMA(0,1,2)}$

```{r}
bk_arima012_fit_cons<- bk_ts3 %>% 
  model(arima012_constant = ARIMA(total_waste ~ 1 + pdq(0,1,2)))

zoo_arima012_fit <- arima(DSNY_BK_zoo_ts, order = 1 + c(0,1,2))
res_arima012 <- zoo_arima012_fit$residuals
acf(res_arima012, lag.max = 36)
pacf(res_arima012, lag.max = 36)
accuracy(bk_arima012_fit_cons)[4]
```

RMSE = 5528.971. Lag 4 in the PACF is still significant, along with the seasonal lag. 

## ${ARIMA(0,1,2)(1,0,0)}$ with constant

${ARIMA(0,1,2)(1,0,0)_{12}}$

```{r}
bk_arima012_100_fit_seasonal_cons <- bk_ts3 %>% 
  model(arima012_100_constant_seasonal = ARIMA(total_waste ~ 1 + 
                                     pdq(0,1,2) + 
                                     PDQ(1,0,0, period = 12)))
        
zoo_arima012_100_seasonal_fit <- arima(DSNY_BK_zoo_ts, 
                          order = 1 + c(0,1,2),
                          seasonal = list(order = c(1,0L,0L), period = 12))
#names(zoo_arima000_fit)
res_arima012_100_seasonal <- zoo_arima012_100_seasonal_fit$residuals
acf(res_arima012_100_seasonal, lag.max = 36)
pacf(res_arima012_100_seasonal, lag.max = 36)
accuracy(bk_arima012_100_fit_seasonal_cons)[4]
```

RMSE = 4427.169

## ${ARIMA(0,1,4)}$ with constant

${ARIMA(0,1,4)}$

```{r}
bk_arima014_fit_cons<- bk_ts3 %>% 
  model(arima014_constant = ARIMA(total_waste ~ 1 + pdq(0,1,4)))

zoo_arima014_fit <- arima(DSNY_BK_zoo_ts, order = 1 + c(0,1,4))
res_arima014 <- zoo_arima014_fit$residuals
acf(res_arima014, lag.max = 36)
pacf(res_arima014, lag.max = 36)
accuracy(bk_arima014_fit_cons)[4]
```

RMSE = 5410.265

## ${ARIMA(0,1,4)(1,0,0)}$ with constant

${ARIMA(0,1,4)(1,0,0)_{12}}$

```{r}
bk_arima014_100_fit_seasonal_cons <- bk_ts3 %>% 
  model(arima014_100_constant_seasonal = ARIMA(total_waste ~ 1 + 
                                     pdq(0,1,4) + 
                                     PDQ(1,0,0, period = 12)))
        
zoo_arima014_100_seasonal_fit <- arima(DSNY_BK_zoo_ts, 
                          order = 1 + c(0,1,4),
                          seasonal = list(order = c(1,0L,0L), period = 12))
#names(zoo_arima000_fit)
res_arima014_100_seasonal <- zoo_arima014_100_seasonal_fit$residuals
acf(res_arima014_100_seasonal, lag.max = 36)
pacf(res_arima014_100_seasonal, lag.max = 36)
accuracy(bk_arima014_100_fit_seasonal_cons)[4]
```

RMSE = 4392.534

## Auto-arima

For our final models, we will look and compare the results of an auto-arima model from the feasts package.

```{r}
bk_auto_arima_fit_cons <- bk_ts3 %>%
  model(stepwise = ARIMA(total_waste),
        search = ARIMA(total_waste,
                       stepwise = FALSE,
                       approximation = FALSE))
accuracy(bk_auto_arima_fit_cons)[1:4]
```

```{r}
bk_auto_arima_fit_cons %>% select(.model = stepwise) %>% report()
print("--------------")
bk_auto_arima_fit_cons %>% select(.model = search) %>% report()
```

## Summary of models

$ARIMA(0,0,4)(1,0,0)_{12}$ has $RMSE = 4444.908$ \
${ARIMA(0,1,2)(1,0,0)_{12}}$ has $RMSE = 4427.169$ \
${ARIMA(0,1,4)(1,0,0)_{12}}$ has $RMSE = 4392.534$ \
${ARIMA(3,0,3)}$ has $RMSE = 5219.698$
